### 2020-09-27 0_author-contribution/itt-emotion-authorcontribution.tex

\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[smallextended,natbib]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof

% For the ORCID icon
\usepackage{scalerel}
\usepackage{tikz}
\usetikzlibrary{svg.path}

\definecolor{orcidlogocol}{HTML}{A6CE39}
\tikzset{
  orcidlogo/.pic={
    \fill[orcidlogocol] svg{M256,128c0,70.7-57.3,128-128,128C57.3,256,0,198.7,0,128C0,57.3,57.3,0,128,0C198.7,0,256,57.3,256,128z};
    \fill[white] svg{M86.3,186.2H70.9V79.1h15.4v48.4V186.2z}
                 svg{M108.9,79.1h41.6c39.6,0,57,28.3,57,53.6c0,27.5-21.5,53.6-56.8,53.6h-41.8V79.1z M124.3,172.4h24.5c34.9,0,42.9-26.5,42.9-39.7c0-21.5-13.7-39.7-43.7-39.7h-23.7V172.4z}
                 svg{M88.7,56.8c0,5.5-4.5,10.1-10.1,10.1c-5.6,0-10.1-4.6-10.1-10.1c0-5.6,4.5-10.1,10.1-10.1C84.2,46.7,88.7,51.3,88.7,56.8z};
  }
}

\newcommand\orcidicon[1]{\href{https://orcid.org/#1}{\mbox{\scalerel*{
\begin{tikzpicture}[yscale=-1,transform shape]
\pic{orcidlogo};
\end{tikzpicture}
}{|}}}}

\usepackage{lineno, hyperref}
% \usepackage[hyperfootnotes=false]{hyperref}
\usepackage{array, pbox}
\usepackage{mathtools}
\usepackage{caption}
\usepackage{graphicx}

% To use chinese characters
\usepackage{CJKutf8}

% To add an appendix page
\usepackage[page, title]{appendix}

% Stick footnotes to bottom of the page
\usepackage[bottom]{footmisc}

% To insert commas into large numbers automatically
\usepackage{siunitx}

% To add hypothesis theorem
\usepackage{ntheorem}
\theoremseparator{:}
\newtheorem{hyp}{Hypothesis}

% To add subhypothesis theorems
\makeatletter
\newcounter{subhyp} 
\let\savedc@hyp\c@hyp
\newenvironment{subhyp}
 {%
  \setcounter{subhyp}{0}%
  \stepcounter{hyp}%
  \edef\saved@hyp{\thehyp}% Save the current value of hyp
  \let\c@hyp\c@subhyp     % Now hyp is subhyp
  \renewcommand{\thehyp}{\saved@hyp\alph{hyp}}%
 }
 {}
\newcommand{\normhyp}{%
  \let\c@hyp\savedc@hyp % revert to the old one
  \renewcommand\thehyp{\arabic{hyp}}%
} 
\makeatother

% To add the envelope symbol in Corresponding Author using % \Letter
\usepackage{marvosym}

% to add superscripts to mark the institute of authors
% patch \maketitle:
\newcommand*{\affaddr}[1]{#1} % No op here. Customize it for different styles.
\newcommand*{\affmark}[1][*]{\textsuperscript{#1}}

% For table formatting
\usepackage{multirow}
\usepackage[normalem]{ulem}
\useunder{\uline}{\ul}{}

% For landscape tables
\usepackage{lscape}

\modulolinenumbers[5]

\journalname{Information Technology \& Tourism}

\bibliographystyle{spbasic}

\begin{document}

\title{Cross-culture differences in tourists faced with Japanese hospitality: A text mining and rank-biased overlap measure study of satisfaction and dissatisfaction factors in Chinese and Western cultures}

\titlerunning{Cross-culture differences in tourists faced with Japanese hospitality: A text ...}

\author{Elisa Claire Alem\'an Carre\'on \protect\affmark[1] \orcidicon{0000-0002-6437-0866} \and
        Hugo Alberto Mendoza Espa\~na                                   \protect\affmark[1] \and
        Hirofumi Nonaka                                                 \protect\affmark[1] \and
        Toru Hiraoka                                                    \protect\affmark[2]
}

\authorrunning{E. Alemán Carreón et al.}

\institute{\Letter \hspace*{0.16em} Elisa Claire Alem\'an Carre\'on  \at
            % Nagaoka University of Technology, Nagaoka, Japan \\
            \hspace*{1em} \email{elisa.claire.aleman.carreon@gmail.com} \\
            \hspace*{1em} ORCID: 0000-0002-6437-0866 %\\
            % Corresponding Author %  \\
            % \emph{Present address:P.C. 940-2033, Ribbon Nagaoka B104, 1128-3 Kaminozoki-machi, Nagaoka, Niigata, Japan}
        \and
            \hspace*{1em} Hugo Alberto Mendoza Espa\~na \at
            % Nagaoka University of Technology, Nagaoka, Japan \\
            \hspace*{1em} \email{mendoza.espana@gmail.com}
        \and
            \hspace*{1em} Hirofumi Nonaka \at
            % Nagaoka University of Technology, Nagaoka, Japan \\
            \hspace*{1em} \email{nonaka@kjs.nagaokaut.ac.jp}
        \and
            \hspace*{1em} Toru Hiraoka \at
            % University of Nagasaki, Nagasaki, Japan \\
            \hspace*{1em} \email{hiraoka@sun.ac.jp}
        \and
            \at \affaddr{\affmark[1] \hspace*{0.15em} Nagaoka University of Technology, Nagaoka, Japan}
        \and
            \at \affaddr{\affmark[2] \hspace*{0.15em} University of Nagasaki, Nagasaki, Japan}
}

\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

\maketitle

\section*{Author Contribution}

H. Nonaka and T. Hiraoka conceived of the presented idea. H. Nonaka and E. Alemán Carreón developed the theory. All members participated in planning the experiments. E. Alemán Carreón and H. Mendoza España performed data collection, the experiments, and computations. E. Alemán Carreón wrote the manuscript with support from H. Nonaka and H. Mendoza España. E. Alemán Carreón and H. Nonaka participated in the discussion of results. T. Hiraoka supervised the project.

\end{document}