### 2020-09-27 0_coverletter/V2_first revision/itt-emotion-coverletter.tex

\documentclass{letter}
\makeatletter
  \def\@texttop{}
\makeatother

\signature{Elisa Claire Alemán Carreón}
\address{
\textbf{Corresponding Author:}\\
Elisa Claire Alemán Carreón \\
Nagaoka University of Technology, Nagaoka, Japan \\
P.C. 940-2033, Ribbon Nagaoka B104, \\
1128-3 Kaminozoki-machi, Nagaoka, Niigata, Japan \\
E-mail: elisa.claire.aleman.carreon@gmail.com
}

\addtolength{\headheight}{-11em}
\addtolength{\textheight}{10em}
\addtolength{\textwidth}{8em}
\addtolength{\oddsidemargin}{-6em}

\begin{document}

\begin{letter}{
Information Technology \& Tourism Editorial Board \\
Editor-in-Chief:  Zheng Xiang\\
Virginia Tech, USA}

\opening{Dear Zheng Xiang and ITT Editorial Board:}

We wish to submit a revision for an original research paper JITT-D-19-00095, which is currently under the revised title "Cross-culture differences in tourists faced with Japanese hospitality: A text mining and rank-biased overlap measure study of satisfaction and dissatisfaction factors in Chinese and Western cultures" for further review by Information Technology \& Tourism.

We appreciate the consideration by your journal, the extensive reviews and are grateful for the extension to our deadline that gave us an opportunity to work on this manuscript better under the current COVID-19 pandemic circumstances. This revision required preparing an environment for doing work from home, as well as new experiments done with a data set that is more fair and balanced as pointed out by reviewers. Finally, thanks to our reviewers, our paper is now done under a new focus and research rationale: The Japanese hospitality \textit{Omotenashi} and how different cultures interact with it in terms of satisfaction factors. We found that while Western customers are satisfied with mostly managerial factors as well as Japanese hospitality and service, Chinese customers are more inclined to be satisfied with environmental factors (such as location and transport),  and that they do not prioritize hospitality in their satisfaction while staying at Japanese hotels. We discuss theories on the reason behind this in our revised paper.

We believe our new results and discussion are very interesting, and hope we continue to be in your consideration.

\closing{Sincerely,}

\end{letter}

\end{document}