### 2020-09-28 1_manuscript/Chinese_keys_list.tex
Back on track, copied from recovery pdf from 2020 sep 27 20:00

\begin{CJK}{UTF8}{gbsn}不错\end{CJK} (not bad)
\begin{CJK}{UTF8}{gbsn}大\end{CJK} (big)
\begin{CJK}{UTF8}{gbsn}干净\end{CJK} (clean)
\begin{CJK}{UTF8}{gbsn}早餐\end{CJK} (breakfast)
\begin{CJK}{UTF8}{gbsn}交通\end{CJK} (traffic)
\begin{CJK}{UTF8}{gbsn}棒\end{CJK} (great)
\begin{CJK}{UTF8}{gbsn}近\end{CJK} (near)
\begin{CJK}{UTF8}{gbsn}购物\end{CJK} (shopping)
\begin{CJK}{UTF8}{gbsn}环境\end{CJK} (surroundings)
\begin{CJK}{UTF8}{gbsn}地铁\end{CJK} (subway)
\begin{CJK}{UTF8}{gbsn}卫生\end{CJK} (health)
\begin{CJK}{UTF8}{gbsn}新\end{CJK} (new)
\begin{CJK}{UTF8}{gbsn}推荐\end{CJK} (recommend)
\begin{CJK}{UTF8}{gbsn}选择\end{CJK} (select)
\begin{CJK}{UTF8}{gbsn}地铁站\end{CJK} (subway station)
\begin{CJK}{UTF8}{gbsn}远\end{CJK} (far)
\begin{CJK}{UTF8}{gbsn}附近\end{CJK} (nearby)
\begin{CJK}{UTF8}{gbsn}周边\end{CJK} (surroundings)
\begin{CJK}{UTF8}{gbsn}赞\end{CJK} (awesome)
\begin{CJK}{UTF8}{gbsn}不错 酒店\end{CJK} (nice hotel)
\begin{CJK}{UTF8}{gbsn}大 房间\end{CJK} (big room)
\begin{CJK}{UTF8}{gbsn}干净 房间\end{CJK} (clean room)
\begin{CJK}{UTF8}{gbsn}近 酒店\end{CJK} (near hotel)
\begin{CJK}{UTF8}{gbsn}新 设施\end{CJK} (new facility)
\begin{CJK}{UTF8}{gbsn}棒 酒店\end{CJK} (great hotel)
\begin{CJK}{UTF8}{gbsn}不错 位置\end{CJK} (nice location)
\begin{CJK}{UTF8}{gbsn}大 床\end{CJK} (big bed)
\begin{CJK}{UTF8}{gbsn}干净 酒店\end{CJK} (clean hotel)
\begin{CJK}{UTF8}{gbsn}近 桥\end{CJK} (near bridge)
\begin{CJK}{UTF8}{gbsn}新 酒店\end{CJK} (new hotel)
\begin{CJK}{UTF8}{gbsn}棒 位置\end{CJK} (great position)
\begin{CJK}{UTF8}{gbsn}不错 服务\end{CJK} (nice service)
\begin{CJK}{UTF8}{gbsn}大 酒店\end{CJK} (big hotel)
\begin{CJK}{UTF8}{gbsn}干净 卫生\end{CJK} (clean and hygienic)
\begin{CJK}{UTF8}{gbsn}近 地铁站\end{CJK} (near subway station)
\begin{CJK}{UTF8}{gbsn}新 装修\end{CJK} (new decoration)
\begin{CJK}{UTF8}{gbsn}棒 服务\end{CJK} (great service)
\begin{CJK}{UTF8}{gbsn}不错 环境\end{CJK} (nice environment)
\begin{CJK}{UTF8}{gbsn}大 超市\end{CJK} (big supermarket)
\begin{CJK}{UTF8}{gbsn}干净 环境\end{CJK} (clean environment)
\begin{CJK}{UTF8}{gbsn}近 站\end{CJK} (near station)
\begin{CJK}{UTF8}{gbsn}新 房间\end{CJK} (new room)
\begin{CJK}{UTF8}{gbsn}棒 早餐\end{CJK} (great breakfast)
\begin{CJK}{UTF8}{gbsn}近 地铁\end{CJK} (near subway)
\begin{CJK}{UTF8}{gbsn}不错 早餐\end{CJK} (nice breakfast)
\begin{CJK}{UTF8}{gbsn}大 商场\end{CJK} (big market)
\begin{CJK}{UTF8}{gbsn}干净 总体\end{CJK} (clean overall)
\begin{CJK}{UTF8}{gbsn}棒 房间\end{CJK} (great room)
\begin{CJK}{UTF8}{gbsn}大 社\end{CJK} (big club)
\begin{CJK}{UTF8}{gbsn}棒 水平\end{CJK} (great level)
\begin{CJK}{UTF8}{gbsn}大 空间\end{CJK} (big space)
\begin{CJK}{UTF8}{gbsn}干净 打扫\end{CJK} (clean up)
\begin{CJK}{UTF8}{gbsn}近 车站\end{CJK} (near the station)
\begin{CJK}{UTF8}{gbsn}棒 温泉\end{CJK} (great hot spring)
\begin{CJK}{UTF8}{gbsn}干净 设施\end{CJK} (clean facilities)
\begin{CJK}{UTF8}{gbsn}近 山\end{CJK} (near mountain)
\begin{CJK}{UTF8}{gbsn}干净 服务\end{CJK} (clean service)
\begin{CJK}{UTF8}{gbsn}近 铁\end{CJK} (Kintetsu)
\begin{CJK}{UTF8}{gbsn}新 斋\end{CJK} (new)
\begin{CJK}{UTF8}{gbsn}大 面积\end{CJK} (large area)
\begin{CJK}{UTF8}{gbsn}棒 体验\end{CJK} (great experience)
\begin{CJK}{UTF8}{gbsn}近 市场\end{CJK} (near market)
\begin{CJK}{UTF8}{gbsn}大 卫生间\end{CJK} (big toilet)
\begin{CJK}{UTF8}{gbsn}新 位置\end{CJK} (new location)
\begin{CJK}{UTF8}{gbsn}价格\end{CJK} (price)
\begin{CJK}{UTF8}{gbsn}一般\end{CJK} (general)
\begin{CJK}{UTF8}{gbsn}中文\end{CJK} (Chinese)
\begin{CJK}{UTF8}{gbsn}距离\end{CJK} (distance)
\begin{CJK}{UTF8}{gbsn}地理\end{CJK} (geography)
\begin{CJK}{UTF8}{gbsn}陈旧\end{CJK} (obsolete)
\begin{CJK}{UTF8}{gbsn}老\end{CJK} (old)
\begin{CJK}{UTF8}{gbsn}华人\end{CJK} (Chinese)
\begin{CJK}{UTF8}{gbsn}一般 设施\end{CJK} (general facilities)
\begin{CJK}{UTF8}{gbsn}陈旧 设施\end{CJK} (obsolete facilities)
\begin{CJK}{UTF8}{gbsn}老 酒店\end{CJK} (old hotel)
\begin{CJK}{UTF8}{gbsn}一般 服务\end{CJK} (general service)
\begin{CJK}{UTF8}{gbsn}陈旧 设备\end{CJK} (obsolete equipment)
\begin{CJK}{UTF8}{gbsn}老 设施\end{CJK} (old facility)
\begin{CJK}{UTF8}{gbsn}一般 酒店\end{CJK} (average hotel)
\begin{CJK}{UTF8}{gbsn}陈旧 房间\end{CJK} (outdated room)
\begin{CJK}{UTF8}{gbsn}老 店\end{CJK} (old shop)
\begin{CJK}{UTF8}{gbsn}一般 早餐\end{CJK} (average breakfast)
\begin{CJK}{UTF8}{gbsn}陈旧 酒店\end{CJK} (outdated hotel)
\begin{CJK}{UTF8}{gbsn}老 装修\end{CJK} (old decoration)
\begin{CJK}{UTF8}{gbsn}一般 味道\end{CJK} (general taste)
\begin{CJK}{UTF8}{gbsn}一般 效果\end{CJK} (general effect)
\begin{CJK}{UTF8}{gbsn}一般 位置\end{CJK} (general location)
\begin{CJK}{UTF8}{gbsn}陈旧 家具\end{CJK} (obsolete furniture)
\begin{CJK}{UTF8}{gbsn}老 建筑\end{CJK} (old building)
\begin{CJK}{UTF8}{gbsn}老 外观\end{CJK} (old appearance)
\begin{CJK}{UTF8}{gbsn}一般 卫生\end{CJK} (general hygiene)
\begin{CJK}{UTF8}{gbsn}陈旧 装修\end{CJK} (old decoration)
\begin{CJK}{UTF8}{gbsn}老 房间\end{CJK} (old room)
\begin{CJK}{UTF8}{gbsn}陈旧 感觉\end{CJK} (Stale feeling)
\begin{CJK}{UTF8}{gbsn}一般 房间\end{CJK} (average room)