### 2020-09-29 0_coverletter/itt-emotion-coverletter.tex
Added new cover letter

\documentclass{letter}
\makeatletter
  \def\@texttop{}
\makeatother

\signature{Elisa Claire Alemán Carreón}
\address{
\textbf{Corresponding Author:}\\
Elisa Claire Alemán Carreón \\
Nagaoka University of Technology, Nagaoka, Japan \\
P.C. 940-2033, Ribbon Nagaoka B104, \\
1128-3 Kaminozoki-machi, Nagaoka, Niigata, Japan \\
E-mail: elisa.claire.aleman.carreon@gmail.com
}

\addtolength{\headheight}{-11em}
\addtolength{\textheight}{10em}
\addtolength{\textwidth}{8em}
\addtolength{\oddsidemargin}{-6em}

\begin{document}

\begin{letter}{
Information Technology \& Tourism Editorial Board \\
Editor-in-Chief:  Zheng Xiang\\
Virginia Tech, USA}

\opening{Dear Zheng Xiang and ITT Editorial Board:}

We wish to submit a revision for an original research paper JITT-D-19-00095, which is currently under the revised title "Cross-culture differences in tourists faced with Japanese hospitality: A text mining and natural language processing study of satisfaction and dissatisfaction factors in Chinese and Western cultures" for further review by Information Technology \& Tourism.

We appreciate the consideration by your journal, the extensive reviews and are grateful for the opportunity to revise this further. This revision required adding new experiments and methodology, namely the introduction of a dependency parser and a part of speech tagger to extract adjective+noun pairs, in order to understand what customers are praising or describing. Finally, thanks to our reviewers, our paper has now another layer of depth by analyzing differences in satisfaction factors across different price ranges for hotels. We found that while Western customers are satisfied with mostly soft factors as well as Japanese hospitality and service, Chinese customers are more inclined to be satisfied with hard factors (such as location, transport, size of the room), and that while they are satisfied with hospitality, in exception of a lack of Chinese language service, they do prioritize hospitality in their satisfaction while staying at Japanese hotels. We discuss theories on the reason behind this in our revised paper. We also found that Chinese tourists in Japan are interested in shopping at all price ranges of hotels, while Western tourists are more inclined to shop when staying in middle to high-priced hotels.

We believe our new results and discussion are very interesting, and hope we continue to be in your consideration.

\closing{Sincerely,}

\end{letter}

\end{document}