### 2020-10-07 3_review-responses/V1_review_responses/olas_review_response.txt

From these two rst ranking keywords (\big" and \clean"), we can assert
that room quality is the most important satisfaction factor for Chinese customers.
This means that Chinese tourists have an expectation of spaciousness
18 E. Aleman Carreon et al.
and cleanliness when coming to Japan, be it by reputation, previous experiences,
or cultural backgrounds.
Ola Jajus

so you went and abandoned the search for discussing it further

kinda like ok so this is happening and the reason is whatever now lets move on
Eli ❤

Oh

Ok I still don’t get where

But you have the same impression then
Ola Jajus
My furry fox girlfriend ❤

had a team of Chinese speaking lab-mates point out the
usage of the word

lab members more professional
Eli ❤

Right
Ola Jajus
My furry fox girlfriend ❤

One key component we found in Chinese customer satisfaction factors is
the inclusion of breakfast within the hotel. This can be inferred from the high
555 frequency with which this keyword was included in the sentences emotionally
classied as positive. While other food-related words were extracted, most of
them were general in nature, like \food" or \eating", and in a lower ranking.
In contrast, the word \breakfast", which is referring to a specic time and
very possibly its inclusion in the hotel commodities, was very frequently used
560 in positive texts compared to other food-related words.
Ola Jajus
My furry fox girlfriend ❤

ok so you explained it all nicely

but you just stopped there without giving further tips to the hotel

i mean ofc the tip is provide free breakfast

but maybe refer to idk some hotel strategy?

if u dont know hotel stuff its hard ik
Eli ❤

Hmmm ok i guess it does come across as withholding information even tho I thought it was obvious
Ola Jajus
My furry fox girlfriend ❤

it is
Eli ❤

Hmm well thanks honey
Ola Jajus
My furry fox girlfriend ❤

ik u probably sum it up later on
Eli ❤

I guess you opened my eyes to why they said that
Ola Jajus
My furry fox girlfriend ❤

yeah maybe u have to translate the results to clear strategies

or end every section of discussion with "Strategies to be implemented: ... include breakfast... ...blah...."
Ola Jajus
My furry fox girlfriend ❤

Moreover, mainland Chinese tourists have expectations about the treatment
towards the Chinese visitors that aren't being met. Be it Chinese language
pamphlets, Chinese texts on instructions for the hotel room and it's
appliances and features (T.V. channels, Wi-Fi setup, etc), or just the treat565
ment towards Chinese people, it seems that their expectations are not being
met. This is natural, since traveling in a strange land without knowledge of
the language can be a daunting experience. Ryan and Mo (2001) also found
that communication diculties was one of the main reasons Chinese customers
would give for not visiting again.
Ola Jajus
My furry fox girlfriend ❤

AGAIN

you got so much to elaborate upon here

even from our experience

ok so they got language bareeer problems SO hotels have to make sure they need to provide a more welcoming environment, hire chinese speaking staff, maybe provide chinese news channels on tv

you could touch on the large no of chinese ryugakusei everywhere in japan and how they could be employed in hotels to increase satisfaction lvls

ik its obvious cuz we both in japan and ryugakusei but even for japanese hotel industry its not crystal clear, and for US ppl its not clear at all

This is all very
Analyzing dierences in satisfaction and dissatisfaction between Chinese and ... 19
dierent from Westerner experiences and while it could be a culture shock to
some, it is mostly seen positively.

this one sounds not professional

"This is all very different" part change to sth like "This lies in striking contrast to" or "This significantly differs with", "This is usually outside the scope of"

also i know you know all this stuff from ur experience but you are not referencing this cultural stuff like sento and bowing and idk is that ok for sure?

"After all, Japanese sta treats all customers
in this respectful manner, but for some customers, this could very well be the
best they have been treated until that moment." this sentence is awkward and too informal

"We can also see that Kozak
585 (2002) and Shanka and Taylor (2004) have also found" also also

you overuse "we can see", change it to sth like "the experimental data suggests", "it can be inferred"

"We can also see that Kozak
585 (2002) and Shanka and Taylor (2004) have also found that hospitality and
sta friendliness is an important determinant in the satisfaction of Western
tourists.
However, we can see"
Eli ❤

ok but the english i told you you can miss because im gonna redo the grammar check
Ola Jajus
My furry fox girlfriend ❤

"we can see from the rest of the keywords" looking at the remaining keywords, we can infer

is gramarly gonna fix informal language or very vague repetitions?

your language is just normal in some places and formal in some; is it gonna detect the discrepancy?

"have found many places unacceptable to their standards" is that an expression for sure? maybe just unacceptable
Ola Jajus
My furry fox girlfriend ❤

"In comparison, our research uncovered that it is mostly linked
to dissatisfaction and that westerners could be said to have high standards of
hygiene when compared to their Chinese counterparts" careful here someone might get offended. maybe use higher standards of room cleanliness; specify its ROOM
Ola Jajus
My furry fox girlfriend ❤

"Some cities in Japan (e.g.
605 Osaka) are currently going under a large number of renovations" ahahah experience
😆
1

"Customers staying in places in renovation
or near a construction" add site

oh man so much english mud

you better fix it all

"Another keyword there is \slow", which upon inspection of example reviews,
610 we concluded that it re
ects the speed of the Internet connection in the hotel
rooms." ORDER screwed up
Eli ❤

order?
Ola Jajus
My furry fox girlfriend ❤

"An interesting detail of the keywords of satisfaction and dissatisfaction,
is that the word \comfortable" is high on the list of satisfaction factors and
\uncomfortable" high on the dissatisfaction factors. It seems that in general,
615 western tourists are highly sensitive to comfort levels in the hotels, and whether
it reaches their expectations." ok so for example what
Eli ❤

ah must be from the new experiment
Ola Jajus
My furry fox girlfriend ❤

thats so curt

it sucks

like comfort of what??? should the hotels invest in beds? balcony sets? big baths?

translate your research to applications
Ola Jajus
My furry fox girlfriend <3 replied to you

order?
no sentence order is shit
Eli ❤

YEAH OK

i told you to leave the english alone jeez

i know it sucks

I'm gonna fix it afterwards
Ola Jajus
My furry fox girlfriend ❤

why didnt you fix it before submitting manuscript?
Eli ❤

because.... that was before i bought grammarly and well apparently it helps a lot

beause apparently my english sucks

now i have it and i plan to use it
👍
1
Ola Jajus
My furry fox girlfriend ❤

"In general, Japan is an expensive place
to visit, which could have an impact in this placement in the ranking." maybe give a frame? like compare a price of room in china and japan in usd?
👍
1

"7.4 Location, location, location" are you sure abt this header?
Eli ❤

well... actually no reviewers complained and it's apparently a phrase used a lot in real estate
👍
1

so its not like... out of nowhere
Ola Jajus
My furry fox girlfriend ❤

"availability of nearby shops" maybe change to proximite to shops

"Upon inspection of examples from the data, we found that most customers
635 were satised if the hotel was nearby to at least two other subjects: subway
or train, and convenience stores." what do you mean by other subjects?

like why other
Eli ❤

oh hmmm... idk really, it can go
Ola Jajus
My furry fox girlfriend ❤

"Japan is a country with a particular public transport system. The rush
hour makes for a subway lled to the brim with people in suits making their
commute, and trains and subway stations in Tokyo create a confusing public
640 transport map for an outside observer. Buses are also available, although less
used than the rail systems in the big cities." particular to peculiar, outside observer to visitor
Ola Jajus
My furry fox girlfriend ❤

"Taxis in Japan
are a luxury compared to other countries. Where in less developed countries
645 the taxi is the cheap method of transport of choice, in Japan, taxis are made
to provide a high-quality experience, with a matching price." ok maybe give a price in usd of train from ueno to tokyo station and taxi so that they get an idea? also if u compare taxi price to usa its comparable so maybe you should add that japanese taxis are expensive compared to other southeast asia countries?
Ola Jajus
My furry fox girlfriend ❤

"This means that
for tourists, subway availability and a good map or plan to travel the city are
of utmost necessity" yo come on dont be out of this century, mention travel apps or google maps or sth

"Japanese convenience stores, on the other hand, are also famous worldwide
650 for their convenience" xDDDDDDD

you dont say

can you rephrase it
Eli ❤

i mean

but they are
😆
1

but yeah sure
Ola Jajus
My furry fox girlfriend ❤

"If some trouble occurred, or a traveler forgot to pack
a certain item, it is almost sure that they can nd it here." hmm i think not,
Ola Jajus
My furry fox girlfriend ❤

"Therefore, considering both transport systems and nearby shops are points
of interest for Chinese and Western tourists, Japanese hotels have to choose
carefully their location from the moment that they are constructed. While not
a top priority, this is a universal factor for both customer groups and it can
be an instant way to generate positive reviews." YES thank you this is what we all wanted
Ola Jajus
My furry fox girlfriend ❤

also more like from the moment they purchase/lease land
👍
1

you could also mention they could purchase land and start buuding after the plans to make new metro line/connection hub are made

like to elaborate a lil on the strategies to take

"After all, even if it doesn't bother some customers, the lack of smell would not
bother any existing customers." i know what ur saying but make it clearer
👍
1

hey im not sure if its a problem but u got "undened 30%" maybe you should comment what that could be and how u could analyze it in future work?
Eli ❤

hmm, well for now it wasnt a problem and i dont wanna angry the reviewer
👍
1
Ola Jajus
My furry fox girlfriend ❤

"However, if hotels want
to attract more and more customers" informal

yeah so u only start to touch on applications a tiny bit in the last part of 7 discussion

basically u need more i think

"Another limitation is that a
big portion of the Asian tourists coming to Japan are Taiwanese and Korean," but istnt chinese veeeery similar to taiwanese

"This
means our hypothesis 2 is rejected (RBOEXT > 0:5), and that the common
ground within both customer groups is also similar in ranking." maybe mention again which one it is

its not like ppl reading the paper gonna remember all ur hypothesis and im always annoyed when i gotta look back
Ola Jajus
My furry fox girlfriend ❤

"Our results and discussion can be utilized as a guideline for managerial
770 decisions when considering Chinese and Western tourists, and we can observe
their stark dierences, as well as their common attributes. In future work...." ok but specify the useful managerial decisions thats like the very worth of this paper
Ola Jajus
My furry fox girlfriend ❤

you worte the paper abt the topic you dont care about

but make it useful for someone who does
Eli ❤

thanks honey
Ola Jajus
My furry fox girlfriend ❤

ok done nice work on future work

ima make dinner im so hungry
❤
1
You replied to My furry fox girlfriend <3

"This
means our hypothesis 2 is rejected (RBOEXT > 0:5), and that the common
ground within both customer groups is also similar in ranking." maybe mention again which one it is
i mean it does say a bit in there but i guess i can add more
Ola Jajus
My furry fox girlfriend ❤

i know, they should remember it after reading the paper but still

for ppl who skim to conclusions...
Eli ❤