### 2021-03-31 2_revisions/V6_ITT_major_revision/0_coverletter/itt-emotion-coverletter.tex
Add new stuff V6 to revisions history

\documentclass{letter}
\makeatletter
  \def\@texttop{}
\makeatother

\signature{Elisa Claire Alemán Carreón}
\address{
\textbf{Corresponding Author:}\\
Elisa Claire Alemán Carreón \\
Nagaoka University of Technology, Nagaoka, Japan \\
P.C. 940-2033, Ribbon Nagaoka B104, \\
1128-3 Kaminozoki-machi, Nagaoka, Niigata, Japan \\
E-mail: elisa.claire.aleman.carreon@gmail.com
}

\addtolength{\headheight}{-11em}
\addtolength{\textheight}{10em}
\addtolength{\textwidth}{8em}
\addtolength{\oddsidemargin}{-6em}

\begin{document}

\begin{letter}{
Information Technology \& Tourism Editorial Board \\
Editor-in-Chief:  Zheng Xiang\\
Virginia Tech, USA}

\opening{Dear Zheng Xiang and ITT Editorial Board:}

We wish to submit a revision for an original research paper JITT-D-19-00095, which is currently under the title "Differences in Chinese and Western tourists faced with Japanese hospitality: A natural language processing approach" for further review by Information Technology \& Tourism.

We appreciate the consideration by your journal. In this revision, we were asked to make the text more concise and to review the English grammar. We performed a full check to the manuscript to remove sections of text that weren't necessary, made the text more concise and organized a few paragraphs differently, as well as making a few paragraphs belong to a new section in our discussion to avoid going off topic. We also improved on the abstract to make it more concise and appealing.

We believe this was an appropriate response to the major revision, and hope we continue to be in your consideration.

\closing{Sincerely,}

\end{letter}

\end{document}